---
title: "Final Project"
author: "Joohyeok"
date: "2024-04-16"
output: html_document
---

```{r setup, include=TRUE}
library(readxl)
library(ggplot2)
library(GGally)
library(corrplot)
library(psych)

data <- read_excel("Assignment8_data.xlsx", col_names = c("X1", "X2", "Y1", "Y2"))
data

```

a) Initial investigation
```{r}
summary(data)

ggpairs(data,lower=list(continuous = wrap("points", color = "blue", size = 1)),
        diag = list(continuous = "barDiag"))+
  theme_bw()+
  theme(strip.text=element_text(size=12))

boxplot(data, 
        main = "Boxplot of Plant Part Lengths",
        names = c("X1", "X2", "Y1", "Y2"),
        ylab = "Length")

```

b) ---For the X group---
i) Display the relevant sample covariance matrix S
```{r}
xvalue<- data[, 1:2]  
xvalue
cov(xvalue, use="complete.obs")

```

ii) List the eigenvalues and describe the percent contributions to the variance
```{r}
eigen_result<-eigen(cov(xvalue, use="complete.obs"))

eigenvalues <- eigen_result$values

total <- sum(eigenvalues)

percent_variance <- eigenvalues / total * 100

eigenvalues
percent_variance

```

(iv) Give the eigenvectors for the principal components you retain
```{r}
eigenvectors <- eigen_result$vectors

eigenvectors

```

(vi) Display scatter plots of pairs of principal components
```{r}
first_pca_result <- prcomp(xvalue, scale. = FALSE)

plot(first_pca_result$x[, 1], first_pca_result$x[, 2], 
     xlab="First Principal Component", ylab="Second Principal Component")

```

c) ---For the Y group---
(i) Display the relevant sample covariance matrix S
```{r}
Yvalue<- data[, 3:4]  
Yvalue
cov(Yvalue, use="complete.obs")

```

(ii) List the eigenvalues and describe the percent contributions to the variance
```{r}
eigen_Yvalue<-eigen(cov(Yvalue, use="complete.obs"))
eigenvalues_Yvalue <- eigen_Yvalue$values

total_Y <- sum(eigenvalues_Yvalue)

percent_variance_Y <- eigenvalues_Yvalue / total_Y * 100

eigenvalues_Yvalue
percent_variance_Y

```

(iv) Give the eigenvectors for the principal components you retain
```{r}
eigenvectors_Y <- eigen_Yvalue$vectors

eigenvectors_Y

```

(vi) Display scatter plots of pairs of principal components
```{r}
pca_result_Y <- prcomp(Yvalue, scale. = FALSE)

plot(pca_result_Y$x[, 1], pca_result_Y$x[, 2], 
     xlab="First Principal Component", ylab="Second Principal Component")

```

d) For the Entire Dataset 
(i) Display the relevant sample covariance matrix S
```{r}
data1 <- read_excel("Assignment8_data.xlsx", col_names = c("X1", "X2", "X3", "X4"))

xx <- data1[, 1:2]
yy <- data1[, 3:4]
colnames(yy) <- c("X1", "X2")

entire <- rbind(xx, yy)
entire
cov(entire, use="complete.obs")

```

(ii) List the eigenvalues and describe the percent contributions to the variance
```{r}
eigen_entire<-eigen(cov(entire, use="complete.obs"))
eigen_entire
eigenvalues_entire <- eigen_entire$values

total_entire <- sum(eigenvalues_entire)

percent_variance_entire <- eigenvalues_entire / total_entire * 100

eigenvalues_entire
percent_variance_entire

```

(iv) Give the eigenvectors for the principal components you retain
```{r}
eigenvectors <- eigen_entire$vectors

eigenvectors

```

(vi) Display scatter plots of pairs of principal components
```{r}
pca_result_entire <- prcomp(entire, scale. = FALSE)

plot(pca_result_entire$x[, 1], pca_result_entire$x[, 2], 
     xlab="First Principal Component", ylab="Second Principal Component")

```
