---
title: "A2_Q3"
author: "Joohyeok"
date: "2024-02-21"
output: html_document
---

```{r setup, include=TRUE}
library(readxl) 
library(ggplot2)
library(GGally)
library(dplyr)
library(tidyverse)

X <- read_excel("w-nat-track-rec.xlsx",col_names=FALSE)

```

a) Construct Y by eliminating the first column of X and display
```{r}

Y<-X[,-1]

Y

```

b) Set R = corr(Y) and display
```{r}
R<-cor(Y)
R

```

c) Compute det(R) and display
```{r}
det_R<-det(R)
det_R

```
- det(R) indicates important about R. det(R) is not zero which means invertible.

d)
```{r}
b<-matrix(c(1,3,1,2,-1,1,-2), nrow=7,ncol=1)
b

x<-solve(R,b)
x

```

e) Compute the norm
```{r}
Rx <- R %*% x
Rx
norm(Rx - b, type="2") 

```
- The calculated norm value of 2.75e^(-14)is as expected. It is very small value which means the x is very accurate and that Rx substantially matches the given vector..

f) Compute w
```{r}
z<-matrix(c(1,-1,-1,-1,0,2,1), nrow=7,ncol=1)
z
w<-R%*%R%*%R%*%R%*%R%*%z
w

```

g) Compute the projection of z on w
```{r}
projection <- function(a, b) {
  (sum(a * b) / sum(b * b)) * b
}

projection(z,w)


```
