---
title: "A4_Q2"
author: "Joohyeok"
date: "2024-03-19"
output: html_document
---

```{r setup, include=TRUE}
library(readxl) 
library(ggplot2)
library(GGally)
library(dplyr)
library(tidyverse)
library(corrplot)
library(reshape2)
library(readr)
library(stats)

data <- read_excel("fleabeetledata.xlsx", sheet = 1)

```

a) Carry out an initial investigation
```{r}
colnames(data) <- c("Experiment", "x1", "x2", "x3", "x4", "x1", "x2", "x3", "x4")

str(data)
summary(data)

sapply(data, var, na.rm = TRUE)

sapply(data, function(x) sum(is.na(x)))
 
```

b) (i) Display the relevant sample covariance matrix S
```{r}
oleracea<- data[, 2:5]  
oleracea
cov(oleracea, use="complete.obs")

```

(ii) List the eigenvalues and describe the percent contributions to the variance
```{r}
eigen_result<-eigen(cov(oleracea, use="complete.obs"))

eigenvalues <- eigen_result$values

total <- sum(eigenvalues)

percent_variance <- eigenvalues / total * 100

eigenvalues
percent_variance

```

(iii) scree plot
```{r}
plot(eigenvalues, type = "b", pch = 19, frame = FALSE,
     xlab = "Principal Component", ylab = "Eigenvalue",
     main = "Scree Plot")

```

(iv) Give the eigenvectors for the principal components you retain
```{r}
eigenvectors <- eigen_result$vectors

eigenvectors

```

(vi) Using at least the first two principal components, display scatter plots of pairs of principal components
```{r}
oleracea <- na.omit(oleracea)

oleracea <- oleracea[!apply(oleracea, 1, function(x) any(is.infinite(x))), ]

pca_result <- prcomp(oleracea, scale. = FALSE)

plot(pca_result$x[, 1], pca_result$x[, 2], 
     xlab="First Principal Component", ylab="Second Principal Component")

```

c) (i) Display the relevant sample covariance matrix S
```{r}
carduorum<- data[, 6:9]  
carduorum
cov(carduorum, use="complete.obs")

```

(ii) List the eigenvalues and describe the percent contributions to the variance
```{r}
eigen_carduorum<-eigen(cov(carduorum, use="complete.obs"))
eigen_carduorum
eigenvalues_carduorum <- eigen_carduorum$values

total_carduorum <- sum(eigenvalues_carduorum)

percent_variance <- eigenvalues_carduorum / total_carduorum * 100

eigenvalues_carduorum
percent_variance

```

(iii) scree plot
```{r}
plot(eigenvalues_carduorum, type = "b", pch = 19, frame = FALSE,
     xlab = "Principal Component", ylab = "Eigenvalue",
     main = "Scree Plot")

```

(iv) Give the eigenvectors for the principal components you retain
```{r}
eigenvectors <- eigen_carduorum$vectors

eigenvectors

```

(vi) Using at least the first two principal components, display scatter plots of pairs of principal components
```{r}
pca_result_carduorum <- prcomp(carduorum, scale. = FALSE)

plot(pca_result_carduorum$x[, 1], pca_result_carduorum$x[, 2], 
     xlab="First Principal Component", ylab="Second Principal Component")

```

d) (i) Display the relevant sample covariance matrix S
```{r}
entire<-rbind(oleracea,carduorum)
entire
cov(entire, use="complete.obs")

```

(ii) List the eigenvalues and describe the percent contributions to the variance
```{r}
eigen_entire<-eigen(cov(entire, use="complete.obs"))
eigen_entire
eigenvalues_entire <- eigen_entire$values

total_entire <- sum(eigenvalues_entire)

percent_variance_entire <- eigenvalues_entire / total_entire * 100

eigenvalues_entire
percent_variance_entire

```

(iii) scree plot
```{r}
plot(eigenvalues_entire, type = "b", pch = 19, frame = FALSE,
     xlab = "Principal Component", ylab = "Eigenvalue",
     main = "Scree Plot")

```

(iv) Give the eigenvectors for the principal components you retain
```{r}
eigenvectors <- eigen_entire$vectors

eigenvectors

```

(vi) Using at least the first two principal components, display scatter plots of pairs of principal components
```{r}
pca_result_entire <- prcomp(entire, scale. = FALSE)

plot(pca_result_entire$x[, 1], pca_result_entire$x[, 2], 
     xlab="First Principal Component", ylab="Second Principal Component")

```
