---
title: "Q2_Final"
author: "Joohyeok"
date: "2024-02-09"
output: html_document
---

```{r setup, include=TRUE}
library(ggplot2)
library(GGally)
library(readr)

data=read.table("baker_corn_field_data.txt")

colnames(data) <- LETTERS[1:ncol(data)]

corn<-data[,c("B","C","G","H")]

Color<-as.factor(corn$B)

colors <- c("red", "orange", "yellow", "green", "blue", "cyan", "pink", "brown")

```

a) create a 3x3 matrix scatter plot
```{r}
ggpairs(corn[,c("C","G","H")],mapping = aes(color = Color),legend=4,
        diag = list(continuous = "barDiag"))+
  ggtitle("Matrix Scatter Plot for Corn Field Data")+
  theme(plot.title=element_text(hjust=0.5))+
  scale_color_manual(values = colors)

```

b) List groups that appear to be relatively more tightly clustered than other groups

- A correlation value close to 1 or -1 indicates a strong clustering of points

- C vs G plot is 2 row 1 column. Group 1, 3 and 7 looks clustered than other groups
- C vs H plot is 3 row 1 column. Group 6 and 7 looks clustered than other groups
- G vs H plot is 3 row 2 column. Group 7 looks clustered than other groups.

c) List groups that appear to be relatively widely dispersed

- A correlation value close to 0 indicates that the points are widely dispersed

- C vs G plot is 2 row 1 column. Group 5 and 6 looks widely dispersed
- C vs H plot is 3 row 1 column. Group 1 looks widely dispersed
- G vs H plot is 3 row 2 column. Group 3 and 8 looks widely dispersed

