---
title: "A5_Q1"
author: "Joohyeok"
date: "2024-04-03"
output: html_document
---

```{r setup, include=TRUE}
library(readxl)
library(ggplot2)
library(GGally)
library(corrplot)
library(psych)

bull_data <- read_excel("bullsalesdata.xlsx")

```

a) Carry out a shortened initial investigation
```{r}
head(bull_data)

ggpairs(bull_data,lower=list(continuous = wrap("points", color = "blue", size = 1)),
        diag = list(continuous = "barDiag"))+
  theme_bw()+
  theme(strip.text=element_text(size=12))

boxplot(bull_data, 
        las = 2, 
        col = rainbow(ncol(bull_data)))  


```

c) Display the sample correlation matrix R
```{r}
cor_matrix <- cor(bull_data)
cor_matrix

corrplot(cor_matrix, method = "color")

```

d) (i) List the eigenvalues and describe the percent contributions to the variance.
```{r}
result <- prcomp(bull_data, scale. = TRUE)

eigenvalues <- result$sdev^2

percent_contributions <- eigenvalues / sum(eigenvalues) *100

eigenvalues
percent_contributions

```

d) (ii) Scree Plot
```{r}
plot(eigenvalues, type = "b", pch = 19, frame = FALSE, 
     xlab = "Principal Component", ylab = "Eigenvalue", 
     main = "Scree Plot")
abline(h = 1, col = "red", lty = 2)

```

Q2) (a) Display the table of results
```{r}
cor_matrix <- cor(bull_data)

fa_result <- fa(cor_matrix, nfactors = 3, fm = 'pa', rotate = "none")

print(fa_result$loadings, cutoff = 0, sort = TRUE)

commun <- fa_result$communality
commun

vaccounted <- fa_result$Vaccounted
vaccounted

```

Q2) (b) Show the error matrix
```{r}
loadings <- fa_result$loadings
uniquenesses <- fa_result$uniquenesses

predicted_correlation <- loadings %*% t(loadings)
diag(predicted_correlation) <- diag(predicted_correlation) + uniquenesses

error_matrix <- cor_matrix - predicted_correlation

error_norm <- sqrt(sum(error_matrix^2))

error_matrix
error_norm

```

Q2) (c) Show plots of the loading vectors for the first two factors
```{r}
loadings <- fa_result$loadings[, 1:2]
loadings

loading_df <- as.data.frame(loadings)
loading_df$Variable <- rownames(loading_df)

ggplot(loading_df, aes(x = Variable, y = PA1)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(y = "Factor 1 Loading", 
       title = "Plot of the Loading Vector for Factor 1")

ggplot(loading_df, aes(x = Variable, y = PA2)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(y = "Factor 2 Loading", 
       title = "Plot of the Loading Vector for Factor 2")

```

Q2) (d) Show the factor scores plot for the first two factors
```{r}
factor_scores <- factor.scores(bull_data, fa_result$loadings)$scores

scores_df <- as.data.frame(factor_scores)

ggplot(scores_df, aes(x = PA1, y = PA2)) +
  geom_point() +
  theme_minimal() +
  labs(x = "PA 1 Score", y = "PA 2 Score", 
       title = "Factor Scores Plot for the First Two Factors")

```

Q3) (a) Display the table of results
```{r}
fa_varimax <- fa(cor_matrix, nfactors = 3, fm = 'pa', rotate = "varimax")

print(fa_varimax$loadings, cutoff = 0, sort = TRUE)

commun <- fa_varimax$communality
commun

vaccounted <- fa_varimax$Vaccounted
vaccounted

```

Q4) (a) Show the error matrix
```{r}
result_for_error <- fa(cor_matrix, nfactors = 2, fm = 'pa', rotate = "none")

loadings_error <- result_for_error$loadings
uniquenesses_error <- result_for_error$uniquenesses

predicted_correlation_error <- loadings_error %*% t(loadings_error)
diag(predicted_correlation_error) <- diag(predicted_correlation_error) + uniquenesses_error

error_matrix4 <- cor_matrix - predicted_correlation_error

error_norm4 <- sqrt(sum(error_matrix4^2))

error_matrix4
error_norm4

```

############################################################################
Q2) (a) Display the table of results
```{r}

cor_matrix <- cor(bull_data)

fa_result <- fa(cor_matrix, nfactors = 3, fm = 'pa', rotate = "none")

print(fa_result$loadings, cutoff = 0, sort = TRUE)

commun <- fa_result$communality
commun

vaccounted <- fa_result$Vaccounted
vaccounted

cor_matrix <- cor(bull_data)


factor_analysis <- fa(cor_matrix, nfactors = 3, rotate = 'none')

# Extract factor loadings
factor_analysis$loadings



```

