---
title: "A4_Q3"
author: "Joohyeok"
date: "2024-03-20"
output: html_document
---

```{r setup, include=TRUE}
library(readxl) 
library(ggplot2)
library(GGally)
library(dplyr)
library(tidyverse)
library(corrplot)
library(reshape2)
library(readr)
library(stats)

data <- read_excel("malifarmdata.xlsx")
data

```

a) Store the data in matrix X
```{r}
X <- as.matrix(data)

head(X)

```

b) (i) Provide at least two indicators for each of these data that justify this claim
```{r}
summary(X)

boxplot(data$Family, main="Family")
boxplot(data$DistRD, main="Distance to Road")
boxplot(data$Cotton, main="Cotton")
boxplot(data$Maize, main="Maize")
boxplot(data$Sorg, main="Sorghum")
boxplot(data$Millet, main="Millet")
boxplot(data$Bull, main="Bullocks")
boxplot(data$Cattle, main="Cattle")
boxplot(data$Goats, main = "Goats")

```


c) Create a data matrix by removing the outliers
```{r}
outlier_rows <- c(25, 34, 52, 57, 62, 69, 72)

cleaned_data <- data[-outlier_rows, ]

X_tilde <- as.matrix(cleaned_data)

head(X_tilde)

```

d) (i) Give the relevant sample covariance matrix S
```{r}
cov(X, use="complete.obs")

```

(ii) List the eigenvalues and describe the percent contributions to the variance
```{r}
eigen_X<-eigen(cov(X, use="complete.obs"))

eigenvalues_X <- eigen_X$values

total <- sum(eigenvalues_X)

percent_variance_X <- eigenvalues_X / total * 100

eigenvalues_X
percent_variance_X

```

(iii) scree plot
```{r}
plot(eigenvalues_X, type = "b", pch = 19, frame = FALSE,
     xlab = "Principal Component", ylab = "Eigenvalue",
     main = "Scree Plot")

```

(iv) Give the eigenvectors for the principal components you retain
```{r}
eigenvectors_X <- eigen_X$vectors

eigenvectors_X

```

(vi) Using at least the first two principal components, display scatter plots of pairs of principal components
```{r}
pca_result_X <- prcomp(X, scale = FALSE)

plot(pca_result_X$x[, 1], pca_result_X$x[, 2], 
     xlab="First Principal Component", ylab="Second Principal Component")

```

e) (i) Give the relevant sample covariance matrix S
```{r}
cov(X_tilde, use="complete.obs")

```

(ii) List the eigenvalues and describe the percent contributions to the variance
```{r}
eigen_X_tilde<-eigen(cov(X_tilde, use="complete.obs"))

eigenvalues_X_tilde <- eigen_X_tilde$values

total <- sum(eigenvalues_X_tilde)

percent_variance_X_tilde <- eigenvalues_X_tilde / total * 100

eigenvalues_X_tilde
percent_variance_X_tilde

```

(iii) scree plot
```{r}
plot(eigenvalues_X_tilde, type = "b", pch = 19, frame = FALSE,
     xlab = "Principal Component", ylab = "Eigenvalue",
     main = "Scree Plot")

```

(iv) Give the eigenvectors for the principal components you retain
```{r}
eigenvectors_X_tilde <- eigen_X_tilde$vectors

eigenvectors_X_tilde

```

(vi) Using at least the first two principal components, display scatter plots of pairs of principal components
```{r}
pca_result_X_tilde <- prcomp(X_tilde, scale = FALSE)

plot(pca_result_X_tilde$x[, 1], pca_result_X_tilde$x[, 2], 
     xlab="First Principal Component", ylab="Second Principal Component")

```


