---
title: "A1_Q1"
author: "Joohyeok"
date: "2024-02-05"
output: html_document
---

```{r setup, include=TRUE}
library(ggplot2)
library(tidyverse)
library(GGally)
library(readxl)
library(car)

df<-read_excel("cutdown_environment.xlsx")
```

(a) Create a matrix scatter plot for the data.
```{r}
ggpairs(df,colums=c("AAT","AST","AHC","Wind","Evap"),
        lower=list(continuous = wrap("points", color = "blue")))+
  theme_bw()+
  theme(strip.text=element_text(size=12))



```

(b) The correlation between AAT and AST is 0.869. It indicates a strong positive linear relationship which is seem in second row first column plot. Since this value is close to 1, there is a strong dependency.


(c) The correlation between EVAP and AHC is -0.816. It indicates a strong negative linear relationship which is seem in fifth row third column plot. Since this value is close to -1, there is a strong dependency.


(d) The correlation between AAT and WIND is -0.098. It indicates a non-linear relationship which is seem in fourth row first column plot. Since this value is close to 0, there is no dependency.
